Post-Disturbance Tree Community Trajectories in a Neotropical Forest

### 2019-02-15 WholePlotTrajectories.Rmd
abstract

---
title: "Post-Disturbance Tree Community Trajectories in a Neotropical Forest"
author:
  - name: Ariane MIRABEL
    authsuperscript: 1*
  - name: Bruno Herault
    authsuperscript: 2
  - name: Eric Marcon
    authsuperscript: 1
affiliation:
  - affsuperscript: 1
    dptuniv: UMR EcoFoG, AgroParistech, CNRS, Cirad, INRA, Université des Antilles, Université de Guyane.
    address: >
      Campus Agronomique,
      97310 Kourou,
      France.
  - affsuperscript: 2
    dptuniv: INPHB, Institut National Polytechnique Félix Houphoüet-Boigny
    address: >
      Yamoussoukro,
      Ivory Coast.
corrauthor:
    email: ariane.mirabel@ecofog.gf
    url: https://github.com/ArianeMirabel
abstract: 
 Understanding the ecological rules underlying the maintenance of tropical forests biodiversity, structure, functioning and dynamics is urgent to anticipate their fate in the global change context.
 The huge diversity of tropical forests is often assumed to be regularly reshaped by natural disturbance yielding a diversity peak at intermediate intensity. This intermediate disturbance hypothesis (IDH), though, remains debated and the resilience of community taxonomic and functional diversity and composition remains controversial. 
 To disentangle the ecological processes driving community response to disturbance, we analyzed tree community trajectories following a disturbance gradient in a Neotropical forest. We examined community diversity, composition and redundancy trajectories over 30 years, considering both taxonomic and functional trajectories based on 7 leaf, stem and life-history traits. We highlighted the cyclic recovery of community taxonomic and functional composition. While pre-disturbance taxonomic differences were maintained over time, functional composition trajectories were quite similar among communities. The IDH did predict community taxonomic diversity response while functional diversity was enhanced whatever the disturbance intensity. 
 Although consistent, the recovery of community composition, diversity and redundancy remained unachieved after 30 years. This acknowledged the need of decades-long cycles with no disturbance to ensure a complete recovery, and questioned tropical forest community resilience after repeated disturbances.
keywords: [Community Ecology, Community Diversity Determinants, Disturbance Trajectories, Intermediate Disturbance Hypothesis, Mid-term Resilience, Neotropical Forests, Taxonomic and Functional Biodiversity]
preamble: >
  \usepackage{textcomp}
  \DeclareUnicodeCharacter{B0}{\textdegree}
  \usepackage{tabu}
  \usepackage{caption}
  \captionsetup{justification = justified}
  \renewenvironment{table}{\begin{table*}}{\end{table*}\ignorespacesafterend}
  \hyphenation{bio-di-ver-si-ty sap-lings re-or-gan-i-za-tion post-dis-tur-bance dis-tur-bance}
bibliography: references.bib
lang: english
keywordlabel: Keywords
corrauthorlabel: Corresponding author
output:
  bookdown::pdf_book:
    base_format: EcoFoG::article
    keep_tex: TRUE
  bookdown::word_document2: default
  bookdown::gitbook: default
---

@@

Community compositon and diversity post-disturbance trajectories were shaped by the recruitment of pioneers, identical for all communities and disturbance intensity.
However, taxonomic and functional composition trajectories were decoupled: while functional trajectories were similar in the functional space and converged towards the recovery of the same initial state, taxonomic trajectories displayed initial composition differences maintained along time.
This decoupling was explained by a decrease of the functional redundancy which mitigated the functional impact of disturbance.
The disturbance impact was contrasted as well between functoinal and taxonomic diversity.
While the functional trajectories remained similar whenever the disturbance intensity, taxonomic trajectories were markedly different above a threshold of 20-25% AGB removed for which taxonomic richness was maximized and taxonomic eveness remained resilient.
The Intermediate Disturbance Hypothesis was then validated regarding community taxonomic response but inferred regarding the functional one.
Taxonomic diversity was maximized following a disturbance of intermediate intensity, that was the threshold of 25% AGB removed, and at an intermediate time after disturbance when the recruitment of early- and late-successional species were balanced, that was around 25 years after disturbance.
Whenever the disturbance intensity, community resilience (in terms of recovery of the pre-disturbance state) was tangible but required several decades and relied on the lottery, random recruitment of rare species. 
Although resilient, the functional redundancy was lower for more than 30 years after disturbance which cautioned against the risks of infrequent species loss and the persistence of disturbance-specific communities [@Herault2018].

>>>>

---
title: "Post-Disturbance Tree Community Trajectories in a Neotropical Forest"
author:
  - name: Ariane MIRABEL
    authsuperscript: 1*
  - name: Bruno Herault
    authsuperscript: 2
  - name: Eric Marcon
    authsuperscript: 1
affiliation:
  - affsuperscript: 1
    dptuniv: UMR EcoFoG, AgroParistech, CNRS, Cirad, INRA, Université des Antilles, Université de Guyane.
    address: >
      Campus Agronomique,
      97310 Kourou,
      France.
  - affsuperscript: 2
    dptuniv: INPHB, Institut National Polytechnique Félix Houphoüet-Boigny
    address: >
      Yamoussoukro,
      Ivory Coast.
corrauthor:
    email: ariane.mirabel@ecofog.gf
    url: https://github.com/ArianeMirabel
abstract: 
 Understand biodiversity trajectories of tropical forests following disturbance and clarify the link between biodiversity and sitruabnce is urgent to anticipate the impact of the global change context.
 The taxonomic and functional facets of tree communities trajectories following disturbance and the underlying processes remain to be clarified. The explicit impact of disturbance intensity, often formalized with the Intermediate Disturbance Hypothesis (IDH), is controversial as well in tropical forests.
 To clarify post-disturbance trajectories and examine the scope of the IDH in tropical forests, we analyzed tree community response to a disturbance gradient in a Neotropical forest. We examined community diversity, composition and redundancy trajectories over 30 years. We considered taxonomic and functional trajectories, based on 7 leaf, stem and life-history traits. We highlighted a decoupling between taxonomic trajectories, different between communities, and functional trajectories that were identical, which we explained by the variations in community functional redundancy. We besides clarified the scope of the IDH, which did not apply to the functional diversity but corroborated the hump-shaped taxonomic diversity pattern. 
 Although consistent, the recovery of community composition, diversity and redundancy remained unachieved after 30 years. This acknowledged the need of decades-long cycles with no disturbance to ensure a complete recovery, and questioned tropical forest community resilience after repeated disturbances.
keywords: [Community Ecology, Community Diversity Determinants, Disturbance Trajectories, Intermediate Disturbance Hypothesis, Mid-term Resilience, Neotropical Forests, Taxonomic and Functional Biodiversity]
preamble: >
  \usepackage{textcomp}
  \DeclareUnicodeCharacter{B0}{\textdegree}
  \usepackage{tabu}
  \usepackage{caption}
  \captionsetup{justification = justified}
  \renewenvironment{table}{\begin{table*}}{\end{table*}\ignorespacesafterend}
  \hyphenation{bio-di-ver-si-ty sap-lings re-or-gan-i-za-tion post-dis-tur-bance dis-tur-bance}
bibliography: references.bib
lang: english
keywordlabel: Keywords
corrauthorlabel: Corresponding author
output:
  bookdown::pdf_book:
    base_format: EcoFoG::article
    keep_tex: TRUE
  bookdown::word_document2: default
  bookdown::gitbook: default
---

@@

Community post-disturbance trajectories in compositon and diversity were shaped by the recruitment of a determined pool of pioneers identical among communities and disturbance intensity.
Nevertheless, taxonomic and functional composition trajectories were decoupled: while functional trajectories remained similar in the functional space and converged towards the recovery of a comparable initial state, taxonomic trajectories shiwed initial differences in community composition that were maintained along time.
This decoupling was explained by a decrease of the functional redundancy which mitigated the functional impact of disturbance.
Community functional and taxonomic diversity response to disturbance was contrasted as well.
While the functional trajectories remained similar whenever the disturbance intensity, taxonomic trajectories were markedly different above a threshold of 20-25% AGB removed for which taxonomic richness was maximized and taxonomic eveness remained resilient.
The Intermediate Disturbance Hypothesis applied to the linkage of disturbance with the taxonomic diversity, but not with the functional diversity.
Taxonomic diversity was maximized following a disturbance of intermediate intensity, that was the threshold of 25% AGB removed, and at an intermediate time after disturbance, that was around 25 years after disturbance when the recruitment of early- and late-successional species were balanced.
Whenever the disturbance intensity, community resilience (in terms of recovery of the pre-disturbance state) was tangible but required several decades and relied upon the random, lottery recruitment of rare species. 
Although resilient, the functional redundancy was lower for more than 30 years after disturbance which cautioned against the risks of infrequent species loss and the persistence of disturbance-specific communities [@Herault2018].

### 2019-02-15 WholePlotTrajectories.tex
abstract

\Abstract{
Understanding the ecological rules underlying the maintenance of
tropical forests biodiversity, structure, functioning and dynamics is
urgent to anticipate their fate in the global change context. The huge
diversity of tropical forests is often assumed to be regularly reshaped
by natural disturbance yielding a diversity peak at intermediate
intensity. This intermediate disturbance hypothesis (IDH), though,
remains debated and the resilience of community taxonomic and functional
diversity and composition remains controversial. To disentangle the
ecological processes driving community response to disturbance, we
analyzed tree community trajectories following a disturbance gradient in
a Neotropical forest. We examined community diversity, composition and
redundancy trajectories over 30 years, considering both taxonomic and
functional trajectories based on 7 leaf, stem and life-history traits.
We highlighted the cyclic recovery of community taxonomic and functional
composition. While pre-disturbance taxonomic differences were maintained
over time, functional composition trajectories were quite similar among
communities. The IDH did predict community taxonomic diversity response
while functional diversity was enhanced whatever the disturbance
intensity. Although consistent, the recovery of community composition,
diversity and redundancy remained unachieved after 30 years. This
acknowledged the need of decades-long cycles with no disturbance to
ensure a complete recovery, and questioned tropical forest community
resilience after repeated disturbances.
}

@@

Community compositon and diversity post-disturbance trajectories were
shaped by the recruitment of pioneers, identical for all communities and
disturbance intensity. However, taxonomic and functional composition
trajectories were decoupled: while functional trajectories were similar
in the functional space and converged towards the recovery of the same
initial state, taxonomic trajectories displayed initial composition
differences maintained along time. This decoupling was explained by a
decrease of the functional redundancy which mitigated the functional
impact of disturbance. The disturbance impact was contrasted as well
between functoinal and taxonomic diversity. While the functional
trajectories remained similar whenever the disturbance intensity,
taxonomic trajectories were markedly different above a threshold of
20-25\% AGB removed for which taxonomic richness was maximized and
taxonomic eveness remained resilient. The Intermediate Disturbance
Hypothesis was then validated regarding community taxonomic response but
inferred regarding the functional one. Taxonomic diversity was maximized
following a disturbance of intermediate intensity, that was the
threshold of 25\% AGB removed, and at an intermediate time after
disturbance when the recruitment of early- and late-successional species
were balanced, that was around 25 years after disturbance. Whenever the
disturbance intensity, community resilience (in terms of recovery of the
pre-disturbance state) was tangible but required several decades and
relied on the lottery, random recruitment of rare species. Although
resilient, the functional redundancy was lower for more than 30 years
after disturbance which cautioned against the risks of infrequent
species loss and the persistence of disturbance-specific communities
\citep{Herault2018}.

>>>>

\Abstract{
Understand biodiversity trajectories of tropical forests following
disturbance and clarify the link between biodiversity and sitruabnce is
urgent to anticipate the impact of the global change context. The
taxonomic and functional facets of tree communities trajectories
following disturbance and the underlying processes remain to be
clarified. The explicit impact of disturbance intensity, often
formalized with the Intermediate Disturbance Hypothesis (IDH), is
controversial as well in tropical forests. To clarify post-disturbance
trajectories and examine the scope of the IDH in tropical forests, we
analyzed tree community response to a disturbance gradient in a
Neotropical forest. We examined community diversity, composition and
redundancy trajectories over 30 years. We considered taxonomic and
functional trajectories, based on 7 leaf, stem and life-history traits.
We highlighted a decoupling between taxonomic trajectories, different
between communities, and functional trajectories that were identical,
which we explained by the variations in community functional redundancy.
We besides clarified the scope of the IDH, which did not apply to the
functional diversity but corroborated the hump-shaped taxonomic
diversity pattern. Although consistent, the recovery of community
composition, diversity and redundancy remained unachieved after 30
years. This acknowledged the need of decades-long cycles with no
disturbance to ensure a complete recovery, and questioned tropical
forest community resilience after repeated disturbances.
}

@@

Community post-disturbance trajectories in compositon and diversity were
shaped by the recruitment of a determined pool of pioneers identical
among communities and disturbance intensity. Nevertheless, taxonomic and
functional composition trajectories were decoupled: while functional
trajectories remained similar in the functional space and converged
towards the recovery of a comparable initial state, taxonomic
trajectories shiwed initial differences in community composition that
were maintained along time. This decoupling was explained by a decrease
of the functional redundancy which mitigated the functional impact of
disturbance. Community functional and taxonomic diversity response to
disturbance was contrasted as well. While the functional trajectories
remained similar whenever the disturbance intensity, taxonomic
trajectories were markedly different above a threshold of 20-25\% AGB
removed for which taxonomic richness was maximized and taxonomic eveness
remained resilient. The Intermediate Disturbance Hypothesis applied to
the linkage of disturbance with the taxonomic diversity, but not with
the functional diversity. Taxonomic diversity was maximized following a
disturbance of intermediate intensity, that was the threshold of 25\%
AGB removed, and at an intermediate time after disturbance, that was
around 25 years after disturbance when the recruitment of early- and
late-successional species were balanced. Whenever the disturbance
intensity, community resilience (in terms of recovery of the
pre-disturbance state) was tangible but required several decades and
relied upon the random, lottery recruitment of rare species. Although
resilient, the functional redundancy was lower for more than 30 years
after disturbance which cautioned against the risks of infrequent
species loss and the persistence of disturbance-specific communities
\citep{Herault2018}.
